### 2014-07-18 paper/main.tex
added uc talk
fixed conflict

\documentclass[10pt]{article}

@@

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Adaptive shrinkage and the false sign rate}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew Stephens$^{1*}$, 
\\
\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
\\
$\ast$ E-mail: Corresponding mstephens@uchicago.edu
\end{flushleft}

@@

It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values. The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
In genomics, for example, a very common goal
is to compare the mean
expression (activity) level of many genes in two conditions.
Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
measurements are made on only a small number of
samples in each condition - sometimes as few as one
sample in each condition. Thus the error in estimates
of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
in estimates of $\beta_j$ still greater.

Suppose that we measure, with error, a series of ``effects", $\beta_1,\dots,\beta_J$. To take just one concrete 
example, $\beta_j$ could be 
the difference in the mean (log) expression levels of gene $j$ ($j=1,\dots,J$) between 2 conditions. In this case, a measurement of
the effect might be the difference in sample means obtained in the two conditions. We will let $\hat\beta_j$ denote the measured value of $\beta_j$, and assume that
each measurement comes with an associated standard error, 
$s_j$. A key aim here will be to take proper account of the fact that some measurements may be more precise than others: that is, to take proper account of variation in $s_j$ across $j$.

A common goal, particularly in genomic studies, is to identify which $\beta_j$ differ from zero. This is commonly tackled
by first computing an effect size estimate ($\hat\beta_j$) and its standard error ($s_j$), converting this to a $Z$ score ($Z_j=\hat\beta_j/s_j$)
and a corresponding $p$ value ($p_j$), testing $H_j:\beta_j=0$.  Then standard methods (e.g. the qvalue package) can be used
to estimate False Discovery Rates at any given threshold.

The second issue is that directly modeling the p values, say via non-parametric methods, without taking account of their precision, 
can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
The solution is to model beta hat as a convolution of some distribution g and an error component.

@@

The potential for shrinkage-based estimation to
address the multiple comparisons setting 
has been highlighted
previously, including Greenland and Robins (1991),
Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]

@@

 Suppose that we are interested in the values of $n$ ``effects" $\beta_j$ ($j=1,\dots,n$). In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
 We assume that we have obtained data $D_1,\dots,D_n$ that provide independent estimates $\hat\beta_1,\dots,\hat\beta_n$ of these effects,
 with corresponding (estimated) standard errors $s_1,\dots,s_n$. For our purposes, these estimates and standard errors could be obtained from the data
 using any method, provided they (approximately) satisfy 
 \begin{equation} \label{eqn:betahat}
 \hat\beta_j | \beta_j, s_j \sim N(\beta_j, s_j).
 \end{equation}
 [In some settings a $t$ distribution assumption for $\hat\beta_j | s_j, \beta_j$ may be more appropriate, and could also be incorporated into our methods; see Discussion.]
 %(See Discussion for ideas on how our methods may be applicable more generally; for example if, in place of $s_j$, we have a $p$ value testing $\beta_j=0$.)

@@

@@

By modeling the z scores directly under the alternative, rather than the z scores as being the truth plus noise,
maybe that is what is most problematic? Because the resulting distribution of z scores is just not possible.

>>>>

\documentclass[11pt]{article}

@@

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{False Discovery Rates: A New Deal}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew Stephens$^{1*}$, 
\\
\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
\\
$\ast$ E-mail: Corresponding mstephens@uchicago.edu
\end{flushleft}

@@

Since its introduction in 19xx by Benjamini and Hochberg 
\cite{bh:xx}, the ``False Discovery Rate" (FDR) has quickly established itself
as a key concept in modern statistics. 
%, and the primary tool by which most practitioners handle large-scale multiple testing.
Here we consider one setting in which FDR-based methods are widely used:
an experimenter has measured a large numbers of ``effects", each with an associated
measurement precision (standard error), 
and the goal is to identify which of the effects we can declare to be non-zero (with some prescribed level of confidence).
This setting arises commonly in many areas of statistics, and particularly in genomics applications.
For example, in genomics
a common goal is to identify ``differentially-expressed genes" -- that is, genes
whose mean expression (activity) level differs in two conditions.

%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
%measurements are made on only a small number of
%samples in each condition - sometimes as few as one
%sample in each condition. Thus the error in estimates
%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
%in estimates of $\beta_j$ still greater.
%

Many tools have been developed for this problem \cite{}, and several are widely used.
Here we argue that these existing tools have sub-optimal features that can be improved on, at least in some circumstances.
In particular we argue that existing tools i) fail to take proper account of variations in the measurement precision of different effects;
ii) make, implicitly, unrealistic assumptions about the distribution of the non-zero effects. We provide methods
to address these problems, implemented in an R package {\tt ashr} (short for ``Adaptive Shrinkage") and demonstrate that these methods can lead to a substantial reduction in estimated FDRs (or, equivalently,
a substantial increase in the number of discoveries at a given FDR).

Although we motivate our tools by FDR, we are providing generic tools for shrinkage, and
these could be useful in settings such as wavelets.
The potential for shrinkage-based estimation to
address the multiple comparisons problems tackled by the FDR
has been highlighted
previously, including Greenland and Robins (1991),
Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]

@@

@@

 \subsection*{Notation, and existing methods}

Suppose that we are interested in the values of $J$ ``effects" $\beta_j$ ($j=1,\dots,J$). In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
 We assume that we have obtained data $D_1,\dots,D_J$ that provide independent estimates $\hat\beta_1,\dots,\hat\beta_J$ of these effects,
 with corresponding (estimated) standard errors $s_1,\dots,s_J$. For our purposes, these estimates and standard errors could be obtained from the data
 using any method, provided they (approximately) satisfy 
 \begin{equation} \label{eqn:betahat}
 \hat\beta_j | \beta_j, s_j \sim N(\beta_j, s_j).
 \end{equation}
 [In some settings a $t$ distribution assumption for $\hat\beta_j | s_j, \beta_j$ may be more appropriate, and could also be incorporated into our methods; see Discussion.]

A standard genomics analysis pipeline would first convert these numbers to a $Z$ score ($Z_j=\hat\beta_j/s_j$)
and a corresponding $p$ value ($p_j$), testing $H_j:\beta_j=0$.

Definition of FDR and lfdr (Bayesian); connection to frequentist definitions (Storey).

Then standard methods (e.g. the qvalue package) can be used
to estimate False Discovery Rates at any given threshold.

\subsection*{Problems with existing methods}

Illustration (Figure).

Note that the problem could be reduced by discarding measurements that are too noisy. However,
note limitation of this.

The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, without taking account of their precision, 
can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
The solution is to model beta hat as a convolution of some distribution g and an error component.

Illustration (Figure).

 %(See Discussion for ideas on how our methods may be applicable more generally; for example if, in place of $s_j$, we have a $p$ value testing $\beta_j=0$.)

@@

Note: we'd like to make sure it behaves ok in the case where most observations are null.
Perhaps simulate with 999 null and 1 alternative?
Also maybe look at performance for small sample sizes - e.g. n=1?

@@
